---
title: "Caret Pre-Processing"
author: "Gergely Horvath"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

Goal

It is your job to predict the sales price for each house. For each Id in the test set, you must predict the value of the SalePrice variable. 


Metric

Submissions are evaluated on Root-Mean-Squared-Error (RMSE) between the logarithm of the predicted_pls value and the logarithm of the observed sales price. (Taking logs means that errors in predicting expensive houses and cheap houses will affect the result equally.)

# Load Libraries
```{r}
library(earth)
library(shiny)
library(miniUI)
library(ggplot2)
library(data.table)
library(caret)
library(doParallel)
library(outliers)
library(pls)
library(broom)
library(scales)
```

# Set Up Shiny Gadget
```{r}
ggbrush <- function(data, xvar, yvar) {

  ui <- miniPage(
    gadgetTitleBar("Drag to select points"),
    miniContentPanel(
      # The brush="brush" argument means we can listen for
      # brush events on the plot using input$brush.
      plotOutput("plot", height = "100%", brush = "brush")
    )
  )

  server <- function(input, output, session) {

    # Render the plot
    output$plot <- renderPlot({
      # Plot the data with x/y vars indicated by the caller.
      ggplot(data, aes_string(xvar, yvar)) + geom_point()
    })

    # Handle the Done button being pressed.
    observeEvent(input$done, {
      # Return the brushed points. See ?shiny::brushedPoints.
      stopApp(brushedPoints(data, input$brush))
    })
  }

  runGadget(ui, server)
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```


# Import Data

Load data and rbind test set and training set.
```{r}
data_descr <- read.csv(file = "data/data_description.txt")
sample_submission <- read.csv(file = "data/sample_submission.csv")
ds_test <- read.csv(file = "data/test.csv")
ds_test <- as.data.table(ds_test)
ds_test$SalePrice <- NA
ds_test$status <- "test"
ds_train <- read.csv(file = "data/train.csv")
ds_train <- as.data.table(ds_train)
ds_train$status <- "train"
ds_all <- rbind(ds_train, ds_test)
remove(ds_train)
remove(ds_test)
str(ds_all)
```





# Clean Up Training Data

## Remove Predictors with too many NAs
```{r show NA vars to be removed}
trn <- copy(ds_all[status == "train"])
# determine which variables should be kept, i.e. number of NAs low
var_na <- trn[, colMeans(is.na(trn)) < .95]
trn[, -..var_na]
```

```{r remove NA vars}
trn <- trn[, ..var_na]
remove(var_na)
head(trn)
```



## Remove Near-Zero-Variance Predictors

```{r print near zero variance predictors}
near_zero_vars <- nearZeroVar(trn,freqCut = 95/5)
trn[, ..near_zero_vars]
```

```{r remove near zero variance predictors}
trn <- trn[, -c(..near_zero_vars)]
remove(near_zero_vars)
head(trn)
```

## Clean Up NAs in order to get rid of incomplete cases

```{r}
# library(purrr)
# library(dplyr)
# library(tidyr)
# ds_all %>% 
#   mutate(
#     across(where(is.numeric), ~replace_na(.x, 0))
#   ) %>%
#   as.data.table()

# does almost the same, but returns only is.numeric columns
# as.data.table(lapply(keep(ds_all, is.numeric), function(x) replace(x, is.na(x), 0)))

# replace NA with zero in numeric columns
numeric_names <- names(Filter(function(x) is.numeric(x), trn))
trn[, (numeric_names) := lapply(.SD, function(x) replace(x, is.na(x), 0)),
       .SDcols = numeric_names]
remove(numeric_names)

# replace NA with 'na' in character columns
character_names <- names(Filter(function(x) is.character(x), trn))
trn[, (character_names) := lapply(.SD, function(x) replace(x, is.na(x), "na")),
       .SDcols = character_names]
remove(character_names)
```


## Reduce Collinearity

Correlation matrix:
```{r}
# get numeric columns
charidx <- unname(sapply(trn, function(x) is.numeric(x)))
dt_nr <- trn[, ..charidx]

# exclude SalePrice as we only want to analyze the correlation of predictors
dt_nr <- dt_nr[, -c("SalePrice")]

# also exclude TotRmsAbvGrd as it correlates strongly with variable GrLivArea, which we want to keep
dt_nr <- dt_nr[, -c("TotRmsAbvGrd")]
remove(charidx)

correlations <- cor(dt_nr, use = "complete.obs")

corrplot::corrplot(correlations, order = "hclust",tl.cex = 0.5)
```

Identify high correlation variables:
```{r}
highCorr <- findCorrelation(correlations, cutoff = 0.75, names = T)
head(trn[, ..highCorr])
```

Relevance of OverallQual, with which other varibles does it correlate?
```{r}
trn[, table(OverallQual)]
sort(correlations[,"OverallQual"],decreasing = T)
```


Remove high correlation variables:

```{r}
trn <- trn[, -(..highCorr)]
remove(highCorr)
remove(correlations)
trn
```

## Linear Dependencies
```{r}
combo_info <- findLinearCombos(dt_nr)
remove(dt_nr)
combo_info
remove(combo_info)
```



## Feature Selection

```{r}
trn
```



## Create Dummy Variables

```{r}
ds_train_dummies <- predict(dummyVars(" ~ .", data = trn), newdata = trn)

ds_train_dummies <- as.data.frame(ds_train_dummies)

# remove dummy because it does not exist in the test set
ds_train_dummies$HouseStyle2.5Fin <- NULL
ds_train_dummies$GarageQualEx <- NULL
ds_train_dummies$Id <- NULL
ds_train_dummies
```



# TRAIN MODELS

## PLS

### Train Model PLS

```{r}
# introduce a quadratic element of OverallQual as the price of high quality houses is being underestimated strongly by the prediction
# trn$OverallQual2 <- trn$OverallQual^2
# trn[, OverallQual := NULL]

model_pls <-
  plsr(
    formula = SalePrice ~ .,
    # data = trn,
    data = ds_train_dummies,
    scale = T,
    center = T,
    validation = "CV",
    ncomp = 14
  )

summary(model_pls)
```


RMSE is 24,679 (see line 'CV')
Variance explained by model (see line 'X'): 33.57%?



### Plot Model PLS

```{r}
plot(model_pls)
```


### Validation Plot

```{r}
validationplot(model_pls)
```

MSEP plot:
```{r}
validationplot(model_pls, val.type="MSEP")
```


R2 plot:
```{r}
validationplot(model_pls, val.type="R2")
```


### Prediction & ggbrush() of Training Set

```{r}
pred_train_pls <- predict(model_pls, ds_train_dummies, ncomp = 14)
# shiny
# ggbrush(ds_train_dummies, "SalePrice", "pred_train_pls")
```


### Plot Observed vs Prediction

```{r}
plot(x = ds_train_dummies$SalePrice, y = pred_train_pls)
abline(0,1)
```


### Plot Prediction vs Residuals

```{r}
plot(x = model_pls$fitted.values, y = model_pls$residuals)
```


### RMSE

```{r}
rmse_pls <- round(sqrt(mean((pred_train_pls - trn$SalePrice)^2)), 0)
rmse_pls
```


### Prediction Test Set

Convert test data to dummies:

```{r}
ds_test <- ds_all[status == "test", -c("SalePrice")]
vars <- names(ds_test) %in% names(trn)
ds_test <- ds_test[, ..vars]
ds_test[!complete.cases(ds_test)]
remove(vars)
ds_test_dummies <- predict(dummyVars(" ~ .", data = ds_test), newdata = ds_test)
ds_test_dummies <- as.data.frame(ds_test_dummies)
ds_test_dummies
```


Predict:

```{r}
pred_test_pls <- predict(model_pls, ds_test_dummies, ncomp = 14)
ds_train_dummies$Alleyna
ds_test_dummies$Alleyna

# TODO what to do about missing dummy variable in test set?

head(pred_test_pls)
```


## MARS

```{r}
#count unique values for each variable
sapply(lapply(trn, unique), length)
```

### Train Model MARS (earth)
```{r}
ds_train_dummies <- as.data.table(ds_train_dummies)
str(ds_train_dummies)
model_mars_earth <- earth(ds_train_dummies[, -c("SalePrice")], ds_train_dummies$SalePrice)
# predict train
pred_train_mars_earth <- predict(model_mars_earth, ds_train_dummies[, -c("SalePrice")])
pred_train_mars_earth <- as.vector(pred_train_mars_earth)
model_mars_earth
```

### Summary MARS (earth)
```{r}
summary(model_mars_earth)
```

### Plot MARS (earth)
```{r}
plot(model_mars_earth)
```

### Plot Observed vs Prediction MARS (earth)
```{r}
?base::plot
plot(x = I(ds_train_dummies$SalePrice), y = ds_train_dummies$predicted_mars)
abline(0,1)
```


### Plot Prediction vs Residuals

```{r}
plot(x = model_mars_earth$fitted.values, y = model_mars_earth$residuals)
```


### RMSE

```{r}
rmse_mars_earth <- round(sqrt(mean((pred_train_mars_earth - ds_train_dummies$SalePrice)^2)), 0)
rmse_mars_earth
```

### Predict test set
```{r}
pred_test_mars_earth <- predict(model_mars_earth, ds_test_dummies)
pred_test_mars_earth <- as.vector(pred_test_mars_earth)
```



## MARS - CARET

### Train Model

```{r}
indx <- createFolds(ds_train_dummies$SalePrice, returnTrain = TRUE)
ctrl <- trainControl(method = "cv", index = indx)
model_mars_caret <-
  train(
    x = ds_train_dummies[, -c("SalePrice")],
    y = ds_train_dummies$SalePrice,
    method = "earth",
    preProcess = c("BoxCox", "center", "scale"),
    # tuneGrid = expand.grid(degree = 1, nprune = 2:38),
    trControl = ctrl
  )

pred_train_mars_caret <- as.vector(predict(model_mars_caret, ds_train_dummies[, -c("SalePrice", "predicted_pls")]))

model_mars_caret
```

```{r}
summary(model_mars_caret)
```

```{r}
plot(model_mars_caret)
```

### Plot Observed vs Prediction
```{r}
plot(x = ds_train_dummies$SalePrice, y = pred_train_mars_caret)
abline(0,1)
```


### Plot Prediction vs Residuals

```{r}
plot(
  x = pred_train_mars_caret,
  y = (
    pred_train_mars_caret - ds_train_dummies$SalePrice
  )
)
```


### RMSE

```{r}
rmse_mars_caret <- round(sqrt(mean((pred_train_mars_caret - ds_train_dummies$SalePrice)^2)), 0)
rmse_mars_caret
```

### Predict test set
```{r}
pred_test_mars_caret <- predict(model_mars_caret, ds_test_dummies)
pred_test_mars_caret <- as.vector(pred_test_mars_caret)
```


```{r}
marsImp <- varImp(model_mars_caret)
plot(marsImp)
```


## SVM - CARET
```{r}
model_svm_caret <- caret::train(
        SalePrice ~ .,
        data = ds_train_dummies,
        preProcess = c("BoxCox", "center", "scale"),
        method = "svmRadial"
)
model_svm_caret$results
rmse_svm <- min(model_svm_caret$results$RMSE)
```


## RF - CARET

```{r}
# check CPU in terminal: sysctl -a | grep machdep.cpu
cl <- makePSOCKcluster(5)
registerDoParallel(cl)

model_rf_caret <- caret::train(
        SalePrice ~ .,
        data = ds_train_dummies,
        preProcess = c("BoxCox", "center", "scale"),
        method = "rf"
)

stopCluster(cl)
remove(cl)

model_rf_caret$results
rmse_rf <- min(model_rf_caret$results$RMSE)
```


# Submission

```{r}
submission <- data.frame(Id = ds_test$Id,  SalePrice = pred_test_mars_caret)
submission

write.table(
        submission,
        file = "submission.csv",
        row.names = F,
        sep = ",",
        quote = FALSE
)
```


