---
title: "Caret Pre-Processing"
author: "Gergely Horvath"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

Goal

It is your job to predict the sales price for each house. For each Id in the test set, you must predict the value of the SalePrice variable. 


Metric

Submissions are evaluated on Root-Mean-Squared-Error (RMSE) between the logarithm of the predicted value and the logarithm of the observed sales price. (Taking logs means that errors in predicting expensive houses and cheap houses will affect the result equally.)


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(data.table)
library(caret)
library(doParallel)
```

# Import Data
```{r}
data_descr <- read.csv(file = "data/data_description.txt")
sample_submission <- read.csv(file = "data/sample_submission.csv")
ds_test_orig <- read.csv(file = "data/test.csv")

ds_test <- copy(ds_test_orig)

ds_test <- as.data.table(ds_test)
ds_test$SalePrice <- NA
ds_test$status <- "test"
ds_train <- read.csv(file = "data/train.csv")
ds_train <- as.data.table(ds_train)
SalePriceVector <- ds_train$SalePrice
ds_train$status <- "train"
ds_full <- rbind(ds_train, ds_test)
remove(ds_train)
remove(ds_test)
ds_full
```

# Remove Predictors with too many NAs
```{r id}
var_na <- ds_full[, colMeans(is.na(ds_full)) < .9]
ds_full[, -..var_na]
```

```{r exclude too many NAs vars}
ds_full <- ds_full[, ..var_na]
remove(var_na)
ds_full
```

```{r show character vectors and number of unique values in each column}
charidx <- unname(sapply(ds_full, function(x) is.character(x)))
ds_full_char <- ds_full[, ..charidx]
remove(charidx)
apply(ds_full_char, 2, function(x) length(unique(x)))
#count unique values for each variable
sapply(lapply(ds_full, unique), length)
```

```{r ad_hoc}
ds_full_char[, table(Neighborhood)]
```

# Convert to Factors
```{r}
# show number of incomplete cases
nr_incomplete_cases_0 <- nrow(ds_full[!complete.cases(ds_full)])

# determine index of character variables to be converted to factors
conv <- sapply(ds_full, function(x) is.character(x) && length(unique(x)) < 50)

# extract variables to be converted to factors into separate dt
ds_factors <- ds_full[, ..conv]

# replace NAs with 'None'
ds_factors <- lapply(ds_factors, function(x) replace(x, is.na(x), "None"))

# convert character vectors to factor
ds_factors <- lapply(ds_factors, as.factor)

# reformat from list to factors
ds_factors <- as.data.table(ds_factors)

# cbind factors and non-factors
non_factor_names <- setdiff(names(ds_full), names(ds_factors))
ds_full <- ds_full[, ..non_factor_names]
ds_full <- cbind(ds_full, ds_factors)
remove(ds_factors)

nr_incomplete_cases_1 <- nrow(ds_full[!complete.cases(ds_full)])
ds_full[!complete.cases(ds_full)]
```

# Cleaning NAs
```{r}
ds_full[is.na(LotFrontage), LotFrontage := 0]
ds_full[is.na(GarageYrBlt), GarageYrBlt := 0]
ds_full[is.na(MasVnrArea), MasVnrArea := 0]
ds_full[is.na(BsmtFinSF1), BsmtFinSF1 := 0]
ds_full[is.na(BsmtFinSF2), BsmtFinSF2 := 0]
ds_full[is.na(BsmtUnfSF), BsmtUnfSF := 0]
ds_full[is.na(TotalBsmtSF), TotalBsmtSF := 0]
ds_full[is.na(BsmtFullBath), BsmtFullBath := 0]
ds_full[is.na(BsmtHalfBath), BsmtHalfBath := 0]
ds_full[is.na(GarageArea), GarageArea := 0]
ds_full[is.na(GarageCars), GarageCars := 0]
ds_full[is.na(Exterior1st), Exterior1st := "NotAppl"]
ds_full[is.na(Exterior2nd), Exterior2nd := "NotAppl"]
ds_full[is.na(Functional), Functional := "NotAppl"]
ds_full[is.na(SaleType), SaleType := "NotAppl"]
nr_incomplete_cases_2 <- nrow(ds_full[!complete.cases(ds_full[, -"SalePrice"])])
ds_full[!complete.cases(ds_full[, -"SalePrice"])]
```

# Creating Dummy Variables
```{r}
dt <- predict(dummyVars(SalePrice ~ ., data = ds_full), newdata = ds_full)
ds_full <- as.data.table(dt)
remove(dt)
str(ds_full)
```

## Remove Near-Zero-Variance Predictors
Near-Zero-Variance that will be removed from dataset:
```{r}
near_zero_vars <- nearZeroVar(ds_full)
ds_full[, ..near_zero_vars]
```

```{r}
ds_full <- ds_full[, -c(..near_zero_vars)]
remove(near_zero_vars)
head(ds_full)
```



## Reduce Collinearity

Correlation matrix:
```{r}
correlations <- cor(ds_full)
corrplot::corrplot(correlations, order = "hclust",tl.cex = 0.5)
```

Identify high correlation variables:
```{r}
highCorr <- findCorrelation(correlations, cutoff = 0.75)
remove(correlations)
highCorr_names <- names(ds_full[, ..highCorr])
remove(highCorr)
head(ds_full[, ..highCorr_names])
```

Remove high correlation variables:
```{r}
ds_full <- ds_full[, -(..highCorr_names)]
remove(highCorr_names)
ds_full
```

# Linear Dependencies
```{r}
combo_info <- findLinearCombos(ds_full)
combo_info
remove(combo_info)
```


# Split full dataset into Training and Test Set
```{r}
ds_train <- ds_full[status.train == 1]
ds_train[, status.train := NULL]
ds_test <- ds_full[status.train == 0]
ds_test[, status.train := NULL]
```

# Pre-Process
```{r}
set.seed(96)
preproc_alg <- preProcess(ds_train, method = c("BoxCox", "center", "scale", "pca"))
preproc_alg
ds_train_prepr <- predict(preproc_alg, ds_train)
ds_test_prepr <- predict(preproc_alg, ds_test)
```

```{r}
# add back Sales Prices before training the model
ds_train[, SalePrice := SalePriceVector]
ds_train_prepr[, SalePrice := SalePriceVector]
```


# Train Model

## Model LM Caret
```{r}
model_lm_caret <- caret::train(
        SalePrice ~ .,
        data = ds_train_prepr,
        method = "lm"
)
summary(model_lm_caret)
model_lm_caret$results
```

## Model SVM Caret
```{r}
model_svm_caret <- caret::train(
        SalePrice ~ .,
        data = ds_train_prepr,
        method = "svmRadial"
)
model_svm_caret$results
```

## Model RF Caret
```{r}
# check CPU in terminal: sysctl -a | grep machdep.cpu
cl <- makePSOCKcluster(5)
registerDoParallel(cl)

model_rf_caret <- caret::train(
        SalePrice ~ .,
        data = ds_train_prepr,
        method = "rf"
)

stopCluster(cl)
remove(cl)

model_rf_caret$results
```




# Prediction

```{r}
predicted_sale_price <- predict(model_lm_caret, newdata = ds_test_prepr)
predicted_sale_price

submission <- data.frame(Id = ds_test_orig$Id,  SalePrice = predicted_sale_price)
submission

write.table(
        submission,
        file = "submission.csv",
        row.names = F,
        sep = ",",
        quote = FALSE
)
```


