---
title: "House Prices Iowa"
author: "Gergely Horvath"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

Goal

It is your job to predict the sales price for each house. For each Id in the test set, you must predict the value of the SalePrice variable. 


Metric

Submissions are evaluated on Root-Mean-Squared-Error (RMSE) between the logarithm of the predicted value and the logarithm of the observed sales price. (Taking logs means that errors in predicting expensive houses and cheap houses will affect the result equally.)


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(data.table)
library(caret)
```

# Version A

## Import Data
```{r}
data_descr <- read.csv(file = "data/data_description.txt")
sample_submission <- read.csv(file = "data/sample_submission.csv")
ds_test <- read.csv(file = "data/test.csv")
ds_test <- as.data.table(ds_test)
ds_test$SalePrice <- NA
ds_test$status <- "test"
ds_train <- read.csv(file = "data/train.csv")
ds_train <- as.data.table(ds_train)
ds_train$status <- "train"
ds_full <- rbind(ds_train, ds_test)
```

## Remove Predictors with too many NAs
```{r first_try}
var_na <- ds_full[, colMeans(is.na(ds_full)) < .9]
ds_full <- ds_full[, ..var_na]
remove(var_na)
```

# Convert to Factors
```{r}
nrow(ds_full[!complete.cases(ds_full)])
conv <- sapply(ds_full, function(x) is.character(x) && length(unique(x)) < 8)
ds_factors <- ds_full[, ..conv]
ds_factors <- lapply(ds_factors, function(x) replace(x, is.na(x), "None"))
ds_factors <- lapply(ds_factors, as.factor)
ds_factors <- as.data.table(ds_factors)
non_factor_names <- setdiff(names(ds_full), names(ds_factors))
ds_full <- ds_full[, ..non_factor_names]
ds_full <- cbind(ds_full, ds_factors)
remove(ds_factors)
nrow(ds_full[!complete.cases(ds_full)])
ds_full[!complete.cases(ds_full)]
```

# Cleaning NAs
```{r}
ds_full[is.na(LotFrontage), LotFrontage := 0]
ds_full[is.na(GarageYrBlt), GarageYrBlt := 0]
ds_full[is.na(MasVnrArea), MasVnrArea := 0]
ds_full[is.na(BsmtFinSF1), BsmtFinSF1 := 0]
ds_full[is.na(BsmtFinSF2), BsmtFinSF2 := 0]
ds_full[is.na(BsmtUnfSF), BsmtUnfSF := 0]
ds_full[is.na(TotalBsmtSF), TotalBsmtSF := 0]
ds_full[is.na(BsmtFullBath), BsmtFullBath := 0]
ds_full[is.na(BsmtHalfBath), BsmtHalfBath := 0]
ds_full[is.na(GarageArea), GarageArea := 0]
ds_full[is.na(GarageCars), GarageCars := 0]
ds_full[is.na(Exterior1st), Exterior1st := "NotAppl"]
ds_full[is.na(Exterior2nd), Exterior2nd := "NotAppl"]
ds_full[is.na(Functional), Functional := "NotAppl"]
ds_full[is.na(SaleType), SaleType := "NotAppl"]
ds_full[!complete.cases(ds_full[, -"SalePrice"])]
```


## Remove Near-Zero-Variance Predictors
```{r}
near_zero_vars <- nearZeroVar(ds_full)
ds_full[, ..near_zero_vars]
ds_full <- ds_full[, -c(..near_zero_vars)]
remove(near_zero_vars)
head(ds_full)
```

## Reduce Collinearity
```{r}
ds_coll <- ds_full[, -c("SalePrice")]
numeric_vars <- ds_coll[, sapply(ds_coll, is.numeric)]
ds_coll <- ds_coll[, ..numeric_vars]
remove(numeric_vars)
ds_coll[complete.cases(ds_coll)]
correlations <- cor(ds_coll, use = "complete.obs")
corrplot::corrplot(correlations, order = "hclust",tl.cex = 0.5)
highCorr <- findCorrelation(correlations, cutoff = 0.75)
remove(correlations)
highCorr_names <- names(ds_coll[, ..highCorr])
remove(highCorr)
head(ds_coll[, ..highCorr_names])
remove(ds_coll)
# remove high corr vars
ds_full <- ds_full[, -(..highCorr_names)]
remove(highCorr_names)
ds_full <- ds_full[, -c("Id")]
ds_full[is.na(SalePrice)]
ds_full[!complete.cases(ds_full[, -"SalePrice"])]
#count unique values for each variable
sapply(lapply(ds_full, unique), length)
```

```{r}
ds_train <- ds_full[status == "train"]
ds_train[, status := NULL]
ds_test <- ds_full[status == "test"]
ds_test[, status := NULL]
```

# Model LM Base
```{r}
#count unique values for each variable
sapply(lapply(ds_train, unique), length)
model_lm_base <- lm(SalePrice ~ ., data = ds_train)
summary(model_lm_base)
```

# Model LM Caret
```{r}
# model_lm_caret <- caret::train(
#         SalePrice ~ .,
#         data = ds_train,
#         preProc = c("BoxCox", "center", "scale", "pca"),
#         method = "lm"
# )
# ds_train[, table(Exterior1st)]
# ds_train[, table(Exterior1st)]
# str(ds_train$MSZoning)
# levels(ds_train$MSZoning)
```

# rpart
```{r}
model_rpart <- rpart::rpart(SalePrice ~ ., data = ds_train)
summary(model_rpart)
predicted_rpart <- predict(model_rpart, ds_train)
predicted_rpart
plot(ds_train$SalePrice, predicted_rpart)
```


# Understand the Shape of Data
## Histograms
## BoxPlots
# EDA
```{r}
summary(ds_train)
```

# Feature Engineering
# Basic Model Building
## Linear Regression
## Random Forest
## Gradient Boosting
# Model Tuning
# Ensemble Model Building
# Submission